\section{Method}

\paragraph{Entity types.}
We group time-anchored entities into three categories: explicit, implicit, and timeless, based on how their admissible time is determined under our notion of epistemic availability.

We define \emph{explicit} entities as those with a single earliest admissible year at year-level granularity (e.g., a concrete product launch, public announcement, or release date). For such entities, admissibility uncertainty is negligible, and the entity induces a point-mass distribution over time. For example, ``I just bought the first iPhone that was released last week'' is explicit: the admissible time concentrates on the release year of the first iPhone, so the entity-level admissible-time distribution is a delta mass at that year.

We define \emph{implicit} entities as those without a single right admissible year, where admissibility depends on diffuse public salience, gradual adoption, or evolving discourse, and may reasonably vary across observers. For example, ``There has been a lot of talk about global warming on the news'' is implicit: media coverage and public concern intensified over decades, making it difficult to identify a unique earliest admissible year. As a result, the admissible time for an implicit entity is itself distributed over a range of years.

Finally, \emph{timeless} samples contain no time-anchored entities and impose no temporal constraints.

\paragraph{Entity-level admissible-time representation.}
For each extracted entity $e$, we first classify it as explicit or implicit with respect to epistemic availability, and then assign an admissible-time representation:

\begin{itemize}
    \item For an explicit entity, we assign a single admissible year $y_e$, corresponding to a degenerate admissible-time distribution $T_e = y_e$.
    \item For an implicit entity, we assign an admissible-time support interval $[a_e, b_e]$ and, optionally, a discrete probability mass function $p_e(t)$ over years in this interval. The interval defines the support of admissible times, while the probabilities capture uncertainty over when the entity becomes publicly salient.
\end{itemize}

When probabilities are not explicitly specified, the interval alone is retained (e.g., for conservative aggregation). The representation for implicit entities is intentionally flexible: downstream applications may interpret it as inducing a distribution, while our core method does not assume a particular parametric form.

\paragraph{Operational pipeline.}
\Cref{fig:filtering-flow} summarizes our end-to-end implementation. We operationalize the method as a multi-agent pipeline with the following steps:
\begin{enumerate}
    \item \textbf{Entity Extraction Agent:} Extract time-anchored entities from the question and answer bundle.
    \item \textbf{Entity Type Classification Agent:} Classify each extracted entity as explicit or implicit with respect to epistemic availability.
    \item \textbf{Reasoner Agent:} For each entity, assign an initial admissible-time estimate:
    \begin{itemize}
        \item explicit entities receive a best-estimate year,
        \item implicit entities receive an admissible-time interval and, optionally, per-year probabilities.
    \end{itemize}
    \item \textbf{Search Query Agent:} Generate standalone search queries for each entity.
    \item \textbf{Internet Search Agent:} Retrieve external evidence to ground admissible-time estimates.
    \item \textbf{Reasoner/Synthesizer Agent:} Update entity-level admissible-time representations based on retrieved evidence.
\end{enumerate}

\paragraph{Sample-level aggregation.}
Let $T_e$ denote the admissible-time random variable induced by entity $e$. The admissible time of the full sample is defined as
\[
T_{\text{sample}} = \max_e T_e,
\]
corresponding to the earliest year at which all required knowledge could simultaneously be publicly available. This aggregation rule enforces a strict no-leak constraint: the sample is admissible at year $t$ if and only if every referenced entity is admissible by $t$.

Operationally, this definition induces a sample-level admissible-time distribution. In particular, when a sample contains both explicit and implicit entities, probability mass from implicit entities that falls before an explicit entity’s admissible year is clipped to that year, resulting in a spike at the explicit year and a tail over later years.

For conservative filtering and labeling, we collapse the sample-level distribution to a single year by taking the maximum upper bound across entity-level admissible-time supports. However, retaining the full distribution enables alternative downstream policies, such as random bucketing, stochastic cutoff simulation, or soft inclusion weighting for year-bounded training.

\begin{figure}[t]
    \centering
    \begin{tikzpicture}[
        node distance=0.7cm and 1.0cm,
        box/.style={rectangle, rounded corners, draw=black, thick, align=center, text width=0.23\textwidth, minimum height=0.85cm},
        phase/.style={rectangle, draw=none, align=center, font=\bfseries},
        note/.style={rectangle, draw=black, thick, align=left, text width=0.23\textwidth, font=\small},
        arrow/.style={-Latex, thick}
    ]
        \node[box] (input) {Raw sample\\(all text)};
        \node[box, right=of input, xshift=0.4cm] (ground) {Entity grounding};
        \node[box, right=of ground, xshift=0.4cm] (agg) {CI Aggregation\\(rank-quantiles / LLM)};

        \node[phase, above=of input] (phase1) {Ingest};
        \path let \p1 = (phase1), \p2 = (ground) in node[phase] (phase2) at (\x2,\y1) {Evidence};
        \path let \p1 = (phase1), \p2 = (agg) in node[phase] (phase3) at (\x2,\y1) {Aggregation};

        \node[note, below=of ground] (ground_note) {extract entities\\build queries\\search evidence\\form CI\\take upper bounds};
        \node[box, below=of agg] (assign) {Final year\\label};

        \draw[arrow] (input) -- (ground);
        \draw[arrow] ([yshift=10pt]ground.east) -- ([yshift=10pt]agg.west);
        \draw[arrow] ([yshift=-10pt]ground.east) -- ([yshift=-10pt]agg.west);
        \node at ($(ground.east)!0.5!(agg.west)$) {\vdots};
        \draw[arrow] (agg) -- (assign);

        \draw[arrow] (ground) -- (ground_note);
        \draw[arrow] (ground_note) -- (ground);

        \draw[decorate, decoration={brace, amplitude=5pt}] (ground.north west) -- (ground.north east);
        \node[font=\small] at ($(ground.north)+(0,0.35)$) {$\times N$};
    \end{tikzpicture}
    \caption{Filtering pipeline for a single sample.
    Evidence extraction and CI aggregation are repeated for $N$ samples, then merged into a final year label.}
    \label{fig:filtering-flow}
\end{figure}

\paragraph{Evaluation metrics.}
For now, we evaluate metrics on the \emph{implicit} subset, where gold labels are distributions rather than single years. Let $p_x^{\mathrm{gt}}(t)$ be the gold admissible-time distribution and let $y_x^{\min} = \inf\{t : p_x^{\mathrm{gt}}(t) > 0\}$ be its earliest admissible support year.

\paragraph{Unified primary metric.}
To evaluate both point and distribution outputs with one score, we use a leakage-penalized Wasserstein objective:
\[
\text{Score}_{\lambda}(x)
= W_1\!\left(q_x,p_x^{\mathrm{gt}}\right)
+ \lambda \sum_{t<y_x^{\min}} q_x(t),
\]
where
\[
W_1\!\left(q_x,p_x^{\mathrm{gt}}\right)=\sum_t \left|F_{q_x}(t)-F_{p_x^{\mathrm{gt}}}(t)\right|.
\]
The Wasserstein term is interpretable in year units and works uniformly for delta and spread-out distributions; the second term explicitly penalizes leaked mass before the earliest admissible year. We report no-leak accuracy and $\text{Score}_{\lambda}$ as the primary pair.

Because temporal leakage is asymmetric, we also report a hard leakage indicator:
\[
\text{NoLeak}(x)
= \mathbf{1}\!\left[\forall\, t < y_x^{\min},\; q_x(t) = 0\right].
\]
For a single-year prediction $\hat{y}_x$, we treat it as a degenerate distribution $q_x(t)=\mathbf{1}[t=\hat{y}_x]$, so the same definition applies.


\begin{comment}
\section{Method}

\paragraph{Entity types.}
We group entities into three categories: explicit, implicit, and timeless.
We define \emph{explicit} entities as those with a single right admissible year under our notion of epistemic availability (e.g., a concrete product launch year), so their admissible-time uncertainty is negligible at year granularity and can be treated as a point mass.
For example, ``I just bought the first iPhone that was released last week'' is explicit: the admissible time concentrates on the release year of the first iPhone, so $p_x(t)$ is effectively a delta mass on that year.
We define \emph{implicit} entities as those without a single right year, where admissibility depends on diffuse public salience and can vary across observers.
For example, ``There has been a lot of talk about global warming on the news'' is implicit: coverage has intensified over decades, so it is difficult to pinpoint the earliest year when the statement becomes true. As a result, the earliest-admissible time is itself distributed, and our goal is to return a sample from that distribution.
Finally, \emph{timeless} samples contain no time-anchored entity.
%We do not use a database agent because there is no internal database to cross-reference in this pipeline.

\paragraph{Operational pipeline.}
\Cref{fig:filtering-flow} summarizes our end-to-end implementation.
We operationalize the method as a five-step agent pipeline:
\begin{enumerate}
    \item Entity Extraction Agent: extract time-anchored entities from the question and answer bundle.
    \item Reasoner Agent: assign a best-estimate year and a 95\% confidence interval per entity.
    \item Search Query Agent: generate standalone search queries for each entity.
    \item Internet Search Agent: retrieve evidence from search results to ground the estimates.
    \item Reasoner/Synthesizer Agent: update entity confidence intervals based on the evidence and update the overall estimated year using the maximum upper bound across entities.
\end{enumerate}



%\paragraph{Why not GLiNER.}
%We evaluated GLiNER for entity extraction, but it is a poor fit for this task.%
%Given fixed categories, it must both detect the entity span and map it to a category, and empirically it misses entities frequently in our data.
%Because missed entities directly induce leakage, we instead use LLMs for the full stack (entity extraction through year estimation), with search-based evidence grounding to improve reliability.
\end{comment}
